Tool-driven Revolutions in Archaeological Science

### 2018-11-20 analysis/paper/jcaa-paper.Rmd

>>>>

On another level, we'd like to point to the _technicity_ of tools, in our case coding. Aud Sissel Hoel and Iris van der Tuin are drawing on Ernst Cassirer's „Form und Technik“ [-@cassirer_form_1930] and Gilbert Simondon's „Du monde d'existence des objets techniques" [-@simondon_mode_1958] to define technicity as a term to describe that technology, i. e. tools and machines respectively, interact with the human and with nature [@hoel_ontological_2013, 188]. Technology is said to own an instrumental kind of logos (Cassierer, see @hoel_ontological_2013, 194) and in analogy to language it exercises a certain measure of agency in mediating between human, nature and technology. The use of a scripting language might exemplify this quite nicely, as it is not just a technological tool, but quite rightly named "language" as well.

For our paper the concept of technicity is interesting because in contrast to the concept of technological determinism, which puts technology as a force against the human, van der Tuin and Hoel (as well as Cassirer and Simondon) enfold technology into „the human“ without letting it lose its "foreignness" and further develop the notions of „human“ and „nature“ [@hoel_ontological_2013, 190]. Any change in tools therefore will influence how human researchers interact with their objects of interest. Interestingly it is exactly at the point inbetween continously evolving subjects and objects, where van der Tuin and Hoel place technology and the creation of knowledge [@hoel_ontological_2013, 199]. The "productive back-and-forth" between all parties may challenge established notions of culture, but also offers a concept of formation through differing, called "performative correspondence" [@hoel_ontological_2013, 201].

Another point they make is, that the potential of a certain technology cannot be known before its coming-into-being as it will reflect back on its creation ("concretization" in Simondon's terms) as well as interact with other technologies (developed or yet-to-be developed) [@hoel_ontological_2013, 194]. Seeing the influence the introduction of scientific scripting languages had on ecology we therefore can only expect similar trends to evolve in archaeology, but not predict these.
